### 2016-08-17 draft.tex
.

\section{Results} \label{results}
\subsection{Activity Analysis} \label{activ_analys}
Fig. \ref{full_sim_plots_av} shows the resulting dynamics of the population activity, average $NO$ concentration at the neurons' positions and the average firing threshold. Roughly speaking, both homeostatic mechanisms managed to keep the population activity in the desired range of $3~Hz$. However, what might appear to be slightly faster and stronger random fluctuations in the upper three plots of Fig. \ref{full_sim_plots_av}, turn out to be very regular oscillations across all three parameters depicted in the closeup. While the oscillation amplitude undergoes a rather unpredictable time course, the frequency remains at a constant level of $\simeq 1Hz$. Although one might argue that regular oscillations of the $NO$ concentration are not of much concern in terms of the network's structure and behaviour, this is not true for the population activity.
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_hom/rate_NO_th_compl.eps}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_hom/rate_NO_th_closeup.eps}
\caption{Population activity, average $NO$ concentration (averaged over readouts at neurons' positions) and average firing threshold. Canonical homeostasis was used for $0-250s$ , diffusive for $250-500s$. Activity target was $3 ~Hz$.}
\label{full_sim_plots_av}
\end{figure}

A first approach in comparing the network activity before and after altering the homeostatic mechanism is by means of the distribution of firing rates within the network. As Sweeney et al. point out, it has been experimentally observed that firing rates are rather heterogeneous, resulting in a broad distribution. In Fig. \ref{firing_rate_dist_comp} we compare our results for the firing distribution to the results by Sweeney et al.. 
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/firing_rate_dists/firing_rate_dist_compare.eps}
\caption{Left: Simulation results by Sweeney et al.. Right: Simulation results for the SORN.}
\label{firing_rate_dist_comp}
\end{figure}
We observed a similar tendency of broader firing rate distributions for the diffusive homeostasis, though we did not get as heavy tailed statistics. Sweeney et al. found a reciprocal relation between the broadness of the firing rate distribution and the threshold distribution. The strong oscillations of the threshold make it difficult to estimate the threshold distribution in our case. For the plot in Fig. \ref{thresh_dist_comp}, we calculated each neuron's time average threshold for the case of non-diffusive and diffusive homeostasis and used this as a basis for the distribution. In contrast to Sweeney et al., the statistics did not change significantly.
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_hom/threshold_dist_compare.eps}
\caption{Comparison of threshold distributions observed by Sweeney et al. (left) and in the SORN (right).}
\label{thresh_dist_comp}
\end{figure}

Another possibility of characterizing neural activity is to determine the distribution of interspike intervals (ISIs). Using non-diffusive homeostasis, one can observe a Poisson-like spiking behaviour with a stochastic refractory period, which is typically observed in cortical networks (Fig. . This is not what can be observed for the case of diffusive homeostasis: the ISI distribution reveals preference of interspike times in the range of $1s$, whereas interspike times in of $0.1-0.6s$ are less likely than a Poisson process would predict. This deviation matches the frequency of activity oscillations, see Fig. \ref{full_sim_plots_av}.
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/Spike_Stats/ISI_compare_SORN.eps}
\caption{Top: log-plot of ISI-distribution of diff. and non-diff. homeostasis in the SORN. Bottom: CV-distribution for diff and non-diff. homeostasis.}
\label{ISI_compare}
\end{figure}

As a whole, the oscillating activity imposes a undesirable regularity onto the network's activity, see also Fig. \ref{spike_sequence}. While this effect might be tolerable without any external input, it is likely that the oscillations will overlay and obscure network responses onto external drive, presumably making it unusable for any computational task.
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/Spike_Stats/spike_sequence.eps}
\caption{A spike sequence of 10 randomly picked excitatory neurons of the SORN. At $t=250s$, the homeostatic mechanism is switched to the diffusive variant.}
\label{spike_sequence}
\end{figure}

@@

To complete the expression, one has to find an appropriate description of the $r(\theta)$-relation. As a most simple approximation, we applied a linear regression to the data of the average threshold and the corresponding population activity, see Fig. \ref{thresh_r_linfit}. The constant $C$ in equation \eqref{simple_NO_dyn} is then proportional to the slope of the linear fit. In addition, $C$ is proportional to the amount of NO released per spike, proportional to the number of contributing neurons and antiproportional to the total area of the tissue. We can solve and integrate equations \eqref{Ca_dyn} and \eqref{nNOS_dyn} for a single spike, resulting in a total of $\tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / 3$. Therefore, $C=N_{neur}\cdot \tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / (3\cdot L^2)$ and $C\cdot r(\theta)=\theta \cdot \alpha \cdot N_{neur} \cdot \tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / (3\cdot L^2)$, where $\alpha$ is the slope of the linear fit and $L^2$ the area of the tissue.
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_hom/thresh_r_linfit.eps}
\caption{Linear fit of the relation between the firing threshold (averaged over population) and the population activity. $R^2=0.99996$.}
\label{thresh_r_linfit}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_hom/lin_model_osc.eps}
\caption{Simulation results for equations \eqref{simple_NO_dyn} and \eqref{simple_threshold_dyn}.}
\label{lin_mod_osc}
\end{figure}
Fig. \ref{lin_mod_osc} shows the results achieved from the simulation. We observe that noise is capable of maintaining oscillations in an otherwise stable linear system.

@@

\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/power_spec/power_spectrum3}
\caption{Power Spectrum of the firing threshold and analytic result of equation \eqref{power_spec_thresh_r_g}.}
\label{power_spec_thresh_sim_vs_an}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates and thresholds of excitatory population of 400 Neurons (see equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}). For both values of $\alpha$, rates and thresholds approach the same steady state.}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}

@@

If we define
\begin{equation}
\psi_{ij} = \psi_{ij} \equiv \psi (d(\mathbf{x}_{neur,i},\mathbf{x}_{neur,i}))
\label{interact_matrix_elements}
\end{equation}
we can express the condition \eqref{NO_equil_cond} as
\begin{equation}
\sum_j \psi_{ij}\cdot r_j = NO_0
\label{NO_equil_cond_interact_matrix}
\end{equation}
or, as an operator
\begin{align}
\hat{\psi}\mathbf{r} &= NO_0 \mathbf{n} \label{NO_equil_cond_interact_matrix_operator} \\
\mathbf{n}&\equiv (1,1,...,1)
\end{align}

\begin{figure}
\includegraphics[width=\textwidth]{/home/fschubert/Master/plots/diff_rate_solve.png}
\end{figure}

>>>>

\section{Results} \label{results}
\subsection{Activity Analysis} \label{activ_analys}
Fig. \ref{full_sim_plots_av} shows the resulting dynamics of the population activity, average $NO$ concentration at the neurons' positions and the average firing threshold. Roughly speaking, both homeostatic mechanisms managed to keep the population activity in the desired range of $3~Hz$. However, what might appear to be slightly faster and stronger random fluctuations in the upper three plots of Fig. \ref{full_sim_plots_av}, turn out to be very regular oscillations across all three parameters depicted in the closeup. While the oscillation amplitude undergoes a rather unpredictable time course, the frequency remains at a constant level of $\simeq 1Hz$. Although one might argue that regular oscillations of the $NO$ concentration are not of much concern in terms of the network's structure and behaviour, this is not true for the population activity.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_compl.eps}
\includegraphics[width=\textwidth]{../../plots/diff_hom/rate_NO_th_closeup.eps}
\caption{Population activity, average $NO$ concentration (averaged over readouts at neurons' positions) and average firing threshold. Canonical homeostasis was used for $0-250s$ , diffusive for $250-500s$. Activity target was $3 ~Hz$.}
\label{full_sim_plots_av}
\end{figure}

A first approach in comparing the network activity before and after altering the homeostatic mechanism is by means of the distribution of firing rates within the network. As Sweeney et al. point out, it has been experimentally observed that firing rates are rather heterogeneous, resulting in a broad distribution. In Fig. \ref{firing_rate_dist_comp} we compare our results for the firing distribution to the results by Sweeney et al.. 
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/firing_rate_dist_compare.eps}
\caption{Left: Simulation results by Sweeney et al.. Right: Simulation results for the SORN.}
\label{firing_rate_dist_comp}
\end{figure}
We observed a similar tendency of broader firing rate distributions for the diffusive homeostasis, though we did not get as heavy tailed statistics. Sweeney et al. found a reciprocal relation between the broadness of the firing rate distribution and the threshold distribution. The strong oscillations of the threshold make it difficult to estimate the threshold distribution in our case. For the plot in Fig. \ref{thresh_dist_comp}, we calculated each neuron's time average threshold for the case of non-diffusive and diffusive homeostasis and used this as a basis for the distribution. In contrast to Sweeney et al., the statistics did not change significantly.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/threshold_dist_compare.eps}
\caption{Comparison of threshold distributions observed by Sweeney et al. (left) and in the SORN (right).}
\label{thresh_dist_comp}
\end{figure}

Another possibility of characterizing neural activity is to determine the distribution of interspike intervals (ISIs). Using non-diffusive homeostasis, one can observe a Poisson-like spiking behaviour with a stochastic refractory period, which is typically observed in cortical networks (Fig. . This is not what can be observed for the case of diffusive homeostasis: the ISI distribution reveals preference of interspike times in the range of $1s$, whereas interspike times in of $0.1-0.6s$ are less likely than a Poisson process would predict. This deviation matches the frequency of activity oscillations, see Fig. \ref{full_sim_plots_av}.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/ISI_compare_SORN.eps}
\caption{Top: log-plot of ISI-distribution of diff. and non-diff. homeostasis in the SORN. Bottom: CV-distribution for diff and non-diff. homeostasis.}
\label{ISI_compare}
\end{figure}

As a whole, the oscillating activity imposes a undesirable regularity onto the network's activity, see also Fig. \ref{spike_sequence}. While this effect might be tolerable without any external input, it is likely that the oscillations will overlay and obscure network responses onto external drive, presumably making it unusable for any computational task.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/Spike_Stats/spike_sequence.eps}
\caption{A spike sequence of 10 randomly picked excitatory neurons of the SORN. At $t=250s$, the homeostatic mechanism is switched to the diffusive variant.}
\label{spike_sequence}
\end{figure}

@@

To complete the expression, one has to find an appropriate description of the $r(\theta)$-relation. As a most simple approximation, we applied a linear regression to the data of the average threshold and the corresponding population activity, see Fig. \ref{thresh_r_linfit}. The constant $C$ in equation \eqref{simple_NO_dyn} is then proportional to the slope of the linear fit. In addition, $C$ is proportional to the amount of NO released per spike, proportional to the number of contributing neurons and antiproportional to the total area of the tissue. We can solve and integrate equations \eqref{Ca_dyn} and \eqref{nNOS_dyn} for a single spike, resulting in a total of $\tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / 3$. Therefore, $C=N_{neur}\cdot \tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / (3\cdot L^2)$ and $C\cdot r(\theta)=\theta \cdot \alpha \cdot N_{neur} \cdot \tau_{Ca^{2+}}\cdot ln(1+{Ca^{2+}_{spike}}^3) / (3\cdot L^2)$, where $\alpha$ is the slope of the linear fit and $L^2$ the area of the tissue.
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/thresh_r_linfit.eps}
\caption{Linear fit of the relation between the firing threshold (averaged over population) and the population activity. $R^2=0.99996$.}
\label{thresh_r_linfit}
\end{figure}
\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_hom/lin_model_osc.eps}
\caption{Simulation results for equations \eqref{simple_NO_dyn} and \eqref{simple_threshold_dyn}.}
\label{lin_mod_osc}
\end{figure}
Fig. \ref{lin_mod_osc} shows the results achieved from the simulation. We observe that noise is capable of maintaining oscillations in an otherwise stable linear system.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../../plots/power_spec/power_spectrum3}
\caption{Power Spectrum of the firing threshold and analytic result of equation \eqref{power_spec_thresh_r_g}.}
\label{power_spec_thresh_sim_vs_an}
\end{figure}

@@

\begin{figure}
\includegraphics[width=\textwidth]{../../plots/rate_threshold_simpl_hip_alpha/rate_threshold_simpl_hip_alpha_comb}
\caption{Dynamics of rates and thresholds of excitatory population of 400 Neurons (see equations \eqref{dyn_rate_equation} and \eqref{diff_hom_simpl_sweeney}). For both values of $\alpha$, rates and thresholds approach the same steady state.}
\label{dynamics_rate_threshold_dyn_mean_field_sweeney}
\end{figure}

@@

If we define
\begin{equation}
\psi_{ij} = \psi_{ij} \equiv \psi (d(\mathbf{x}_{neur,i},\mathbf{x}_{neur,i}))
\label{interact_matrix_elements}
\end{equation}
we can express the condition \eqref{NO_equil_cond} as
\begin{equation}
\sum_j \psi_{ij}\cdot r_j = NO_0
\label{NO_equil_cond_interact_matrix}
\end{equation}
or, as an operator
\begin{align}
\hat{\psi}\mathbf{r} &= NO_0 \mathbf{n} \label{NO_equil_cond_interact_matrix_operator} \\
\mathbf{n}&\equiv (1,1,...,1)
\end{align}
The problem of finding the steady-state solution of the homeostatic constraint thus reduces to inverting $\hat{\psi}$:
\begin{equation}
\mathbf{r} = NO_0 \hat{\psi}^{-1} \mathbf{n}
\label{NO_euqil_cond_interact_matrix_operator_solve}
\end{equation}

An example of a distribution of rates acquired from this method is shown in fig. \ref{rate_dist_interact_matrix}.

\begin{figure}
\includegraphics[width=\textwidth]{../../plots/diff_rate_solve.png}
\caption{Distribution of the solution of \eqref{NO_euqil_cond_interact_matrix_operator_solve} with an exponential distance dependent approximation in 2d (similar to the exact solution in 1d). For better statistical significance, calculations where carried out for $1600$ random neurons-positions and $50$ trials in total.}
\label{rate_dist_interact_matrix}
\end{figure}

\subsubsection{Results for an Approximation of the Numerical Steady-State Solution}

If one aims for an appropriate expression for the numerical steady-state solution of a single "point"-source with a constant influx, two things should be of importance: first, it should correctly predict the concentration at the location of the source and second, it should be close to the analytic result in regions further away from the source. One possibility is to use the following "trick" to overcome the divergence of the analytic solution at the origin:

\begin{equation}
NO_{approx.}(r) = \frac{1}{\sqrt{\frac{1}{NO_{source,num.}^2} + \frac{1}{NO(r)^2}}}
\label{Numeric_Solution_Expression_Trick}
\end{equation}

At the pole $NO(0)$, $NO_{approx.}(0)$ then takes the value $NO_{source,num.}$. On the other hand, for larger values of $r$, the term $\frac{1}{NO_{source,num.}^2}$ becomes negligible compared to $ \frac{1}{NO(r)^2}$, thus $NO_{approx.}(r) \rightarrow NO(r)$ for $r \rightarrow \infty$.

To find an approximation for $NO_{source,num.}$, one can interpret this value as a mean of the analytic solution across the square area covered by the corresponding grid cell. As an additional simplification, we substitute the necessary integration over the square grid cell by a circular area of equal size around the source. This calculation yields

\begin{equation}
NO_{source,num.} = \frac{1-h\sqrt{\frac{\lambda}{\pi D}} K_1\left( h\sqrt{\frac{\lambda}{\pi D}}\right) }{h^2 \lambda}
\end{equation}

\subsubsection{Mean-Field approximation for the Variance of Firing Rates}

Generally, Mean-Field approximations reduce the calculation of a full set of interacting entities (or particles, in the context of physics) to a single instance interacting with an effective mean of the entire field, especially ignoring the influence of the single instance onto the whole set of entities and thus the field itself.

In the system presented here, a single neuron with a firing rate $\nu$ must fulfil the equation

\begin{equation}
\nu =\frac{NO_0 - \sum \limits_{i} \nu_i \psi(d_i)}{\psi_0}
\label{sample_neuron_firing_rate}
\end{equation}

where $i$ indexes over all "remaining" neurons and $d_i$ refers to the euclidean distance of the sample neuron to "field"-neuron $i$. $\psi_0$ is the factor that - multiplied with $\nu$ - yields the $NO$-concentration at the origin of the sample neuron.

Depending on the constraints that one imposes onto the system, one might look for different properties to be extracted from this equation. Specifically, in the simulation protocol that has been used in all previous simulations, the average firing rate was forced to a certain value by means of adjusting the target concentration $NO_0$. A Mean-Field approximation for the average firing rate therefore is of no relevance in this particular case.

However, one might try to give an estimate for the variance of the firing rate distribution in dependence of the diffusion constant $D$. Interpreting $\nu$ in \eqref{sample_neuron_firing_rate} as a stochastic variable, one finds

\begin{equation}
Var[\nu] = \frac{Var[\sum \limits_{i} \nu_i \psi(d_i)]}{\psi_0^2}
\label{variance_sample_neuron_firing_rate}
\end{equation}
